---
title: "Isabelle"
author: "Isabelle Tooley"
date: "9/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(DataExplorer)
library(corrplot)
library(ggplot2)
library(caret)
library(DMwR)
library(ROSE)
library(dummies)
library(MLmetrics)
```

READ IN DATA
```{r}
cancer_test <- read.csv("test.csv")
cancer_train <- read.csv("train.csv")

#names(cancer_test)[names(cancer_test) == "Id"] <- "SiteNum"
cancer <- bind_rows(train = cancer_train, test = cancer_test, .id = "Set") %>% mutate(Response = as.factor(Response))
```



EDA

1. Summary
```{r}
# summary of data, look for missing values
summary(cancer)
```

2. Missing values exploration
```{r}
# plot of proportions of missing values
plot_missing(cancer)
```

```{r}
nrow(cancer %>% filter(is.na(Consensus) & is.na(PSSM)))
```
Consensus and PSSM each have 870 missing values, and if you subset the data to include rows where both are missing, you get 870 observations.
*Consensus and PSSM are always missing together*
This makes sense because Consensus is built using PSSM, ANN, SVM

Why are some PSSM measurements missing? Any correlation to whether or not it's a cancer site?
```{r}
mean(cancer_train %>% filter(is.na(PSSM)) %>% pull(Response))
mean(is.na(cancer_train %>% filter(Response == 1) %>% pull(PSSM)))
```

3. Correlations
```{r}
# Which variables are related to each other?
plot_correlation(cancer, type="continuous", 
                 cor_args=list(use="pairwise.complete.obs"))
```

```{r}
quant_vars <- cancer %>%
  select(Iupred.score, ANN, PSSM, SVM, Consensus, normalization) %>%
  drop_na()

corrplot(cor(quant_vars), method = "shade", type = "upper", diag = FALSE, addCoef.col = "black", number.cex = .7)
```



DATA CLEANING

1. Create a new variable that acts as a marker for whether or not PSSM was missing. This might or might not be useful.
```{r}
cancer <- cancer %>%
  mutate(missing_PSSM = if_else(is.na(PSSM), 1, 0))
```

2. Imputation of missing PSSM values with SVM* (very strong linear relationship). 
```{r}
# scatterplot of SVM vs PSSM to check linearity
ggplot(data = cancer, aes(SVM, PSSM)) +
  geom_point() +
  geom_smooth(se = FALSE)

# Stochastic linear regression imputation for PSSM
pssm_lm <- lm(PSSM ~ SVM, data = cancer)
pssm_preds <- predict(pssm_lm, newdata = (cancer %>% filter(is.na(PSSM)))) + 
              rnorm(sum(is.na(cancer$PSSM)), 0, sigma(pssm_lm))

cancer <- cancer %>%
  mutate(PSSM = round(replace(PSSM, is.na(PSSM), pssm_preds), digits = 3))

rm(list = "pssm_lm")

ggplot(data = cancer, aes(SVM, PSSM)) +
  geom_point(aes(color = as.factor(missing_PSSM)))
```

3. Imputation of Consensus variable (average of ANN, SVM, PSSM values)
```{r}
cancer <- cancer %>%
  mutate(Consensus = round(replace(Consensus, is.na(Consensus), 
                                   rowMeans(cancer[is.na(Consensus),4:6])), digits = 3))
```

4. Make dummy variables
```{r}
dummies <- dummyVars(Response ~ .-Set, data = cancer)
cancer_dummies <- data.frame(predict(dummies, newdata = cancer)) %>%
  bind_cols(., cancer %>% select(Set, Response))
```

5. Check correlations and missing values again
```{r}
plot_missing(cancer_dummies)
```

```{r}
# drop ANN, PSSM, SVM because of the direct collinearity to Consensus
# if Consensus was dropped, there's still collinearity between all three, so proceed with Consensus
cancer <- cancer_dummies %>%
  select(-c(ANN, PSSM, SVM))

cancer_train <- cancer %>%
  filter(Set == "train") %>%
  select(-Set)

cancer_test <- cancer %>%
  filter(Set == "test") %>%
  select(-c(Set, Response))

plot_correlation(cancer, type="continuous", 
                 cor_args=list(use="pairwise.complete.obs"))
```



MODELING

1. Class imbalance!
```{r}
table(cancer_train$Response)
```

To "fix" class imbalances, we can...
- down sample so that the frequency of the majority class (0) matches the frequency of minority (1)
- up sample with replacement so that the number of 1s matches the number of 0s
- do a combo method which down samples the majority class and simultaneously creates new data points in the minority class

```{r}
# With classProbs = TRUE, the response variable cannot be 1 and 0
cancer_train <- cancer_train %>% 
  mutate(Response = ifelse(Response == 1, "Yes", "No"))
```


```{r}
# gbm_grid <- expand.grid("n.trees" = 1000, 
#                         "interaction.depth" = seq(1, 5, 2), 
#                         "shrinkage" = seq(.005, .075, .005),
#                         "n.minobsinnode" = c(10, 20))

gbm_grid <- expand.grid("n.trees" = 1000, 
                        "interaction.depth" = 1, 
                        "shrinkage" = .005,
                        "n.minobsinnode" = 10)
```


```{r}
down_ctrl <- trainControl(method = "repeatedcv", 
                          number = 10, 
                          repeats = 2,
                          summaryFunction = prSummary,
                          classProbs = TRUE, 
                          sampling = "down")

down_gbm <- train(Response ~ .,  
                  data = (cancer_train %>% select(-SiteNum)),
                  method = "gbm",
                  tuneGrid = gbm_grid,
                  trControl = down_ctrl, 
                  preProcess = c("center", "scale"),
                  metric = "F",
                  verbose = FALSE)

down_gbm
```

```{r}
up_ctrl <- trainControl(method = "repeatedcv", 
                        number = 10, 
                        repeats = 3,
                        summaryFunction = prSummary,
                        classProbs = TRUE, 
                        sampling = "up")

up_gbm <- train(Response ~ .,  
                data = (cancer_train %>% select(-SiteNum)),
                method = "gbm",
                tuneGrid = gbm_grid,
                trControl = up_ctrl, 
                preProcess = c("center", "scale"),
                metric = "F",
                verbose = FALSE)

up_gbm
```

```{r}
rose_ctrl <- trainControl(method = "repeatedcv", 
                          number = 10, 
                          repeats = 3,
                          summaryFunction = prSummary,
                          classProbs = TRUE, 
                          sampling = "rose")

rose_gbm <- train(Response ~ .,  
                    data = (cancer_train %>% select(-SiteNum)),
                    method = "gbm",
                    tuneGrid = gbm_grid,
                    trControl = rose_ctrl, 
                    preProcess = c("center", "scale"),
                    metric = "F",
                    verbose = FALSE)

rose_gbm
```


```{r}
smote_ctrl <- trainControl(method = "repeatedcv", 
                          number = 10, 
                          repeats = 3,
                          summaryFunction = prSummary,
                          classProbs = TRUE, 
                          sampling = "smote")

smote_gbm <- train(Response ~ .,  
                    data = (cancer_train %>% select(-SiteNum)),
                    method = "gbm",
                    tuneGrid = gbm_grid,
                    trControl = smote_ctrl, 
                    preProcess = c("center", "scale"),
                    metric = "F",
                    verbose = FALSE)

smote_gbm
```

